---
title: "Golem Grad Tortoises"
subtitle: "First TidyTuesday Analysis"
toc: false

execute:
  warning: false
  message: false


format:
  html:
    code-fold: true
    code-summary: "R Code"

---

```{r}
library(tidyverse)

clutch_size_cleaned <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-03-03/clutch_size_cleaned.csv')
tortoise_body_condition_cleaned <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2026/2026-03-03/tortoise_body_condition_cleaned.csv')
```


This data was taken from [TidyTuesday's 2026-03-03 project](https://github.com/rfordatascience/tidytuesday/blob/main/data/2026/2026-03-03/readme.md), which was curated by [Novica Nakov, Free Software Macedonia](https://github.com/novica). The data was originally collected by Dragan Arsovski, Xavier Bonnet, Ana Golubović, and Ljiljana Tomović, who presented it in their paper: [Sex Ratio Bias Triggers Demographic Suicide in a Dense Tortoise Population](https://onlinelibrary.wiley.com/doi/10.1111/ele.70296).


**Background**

Over the generations, people have worked hard to build a society that grants men and women the same respect and opportunities, to varying levels of success. This issue transcends species, as the female tortoises in Lake Prespa in North Macedonia struggle with similar societal concerns. Unable to make their communities recognize them for their own accomplishments, a female tortoise is often reduced to just "Hermann."


To paraphrase from the article and introduction on TidyTuesday:
On an especially population-dense island in North Macedonia's Lake Prespa, male Hermann's tortoises vastly outnumber female Hermann's tortoises. The sexually coercive males  injure the female tortoises, and put them at risk of fatal falls. These harassed females are emaciated, tend to produce smaller clutches, and have lower annual survival rates than neighbouring mainland populations.

Data about the tortoises were taken from two locations on Golem Grad Island: the beach and the plateau. Data was also collected on the nearby mainland population of Hermann's tortoises in Konjsko as a control.

**Masses of Tortoises**

The following plot shows the spread of masses of female vs male tortoises in the three different locations. From this plot, it is clear that the female tortoises in Konjsko tend to be heavier than those on the beach and on the plateau of Golem Grad, though the males at the three locations are comparable in size.

```{r}
ggplot(tortoise_body_condition_cleaned, aes(x = sex, y = body_mass_grams, color = sex)) +
  geom_boxplot() + facet_wrap(vars(locality)) + theme_minimal() +labs(title = "Mass of Females vs Males At Different Localities", x = "Sex and Locality", y = "Body Mass (grams)") + theme(legend.position = "none") 
```


**Body Conditions of Tortoises**

This plot shows the spread of the body condition index of the female and male tortoises in the three different locations. Body condition index refers to a tortoise's mass scaled by its size. As stated by the original article, "Among the genus Testudo, the body condition of females is higher than that of males (Bonnet et al. 2010)". 
However, from this plot, this is only true for the control group in Konjsko. The female tortoises in Golem Grad have similar BCIs to the males.


```{r}
ggplot(tortoise_body_condition_cleaned, aes(x = sex, y = body_condition_index, color = sex)) +
  geom_boxplot() + facet_wrap(vars(locality)) + theme_minimal() +labs(title = "Body Condition of Females vs Males At Different Localities", x = "Sex and Locality", y = "Body Condition Index") + theme(legend.position = "none") 
```




